---
title: "Untitled"
author: "Michael"
date: "21/7/2020"
output:
  pdf_document: default
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

Commit containing full submission

1 Code for reading in the dataset and/or processing the data
2 Histogram of the total number of steps taken each day
3 Mean and median number of steps taken each day
4 Time series plot of the average number of steps taken
5 The 5-minute interval that, on average, contains the maximum number of steps
6 Code to describe and show a strategy for imputing missing data
7 Histogram of the total number of steps taken each day after missing values are imputed
8 Panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends
9 All of the R code needed to reproduce the results (numbers, plots, etc.) in the report

Start


1 Código para leer en el conjunto de datos y / o procesar los datos
```{r  echo = TRUE}
library(ggplot2)  
  activity <- read.csv("./activity.csv")
  activity$date <- as.POSIXct(activity$date,format= "%Y-%m-%d")
  activity_weekday <- weekdays(activity$date)
  activity <- cbind(activity,activity_weekday)
  
  summary(activity)
```
 
2 Histogram of the total number of steps taken each day

```{r echo = TRUE}
total_steps <- with(activity, aggregate(steps, by = list(date), FUN = sum, na.rm = TRUE))
names(total_steps) <- c("date", "steps")
hist(total_steps$steps, main = "Total of steps taken for day",
                      xlab = "Total steps taken for day", 
                      col = "darkblue", ylim = c(0,20), 
                      breaks = seq(0,25000, by=2500))
```
3 Mean and median number of steps taken each day
```{r echo = TRUE}
#Mean of the total number of steps taken for day is:
mean(total_steps$steps)
# 9354.23
#Median of the total number of steps taken for day is:
median(total_steps$steps)
#10395
```

4 Time series plot of the average number of steps taken
```{r echo = TRUE}
#What is the average daily activity pattern?

daily_activity <- aggregate(activity$steps, by=list(activity$interval), FUN=mean, na.rm=TRUE)
names(daily_activity) <- c("interval", "mean")
plot(daily_activity$interval,
     daily_activity$mean, type = "l", 
     col="darkred", lwd = 2, xlab="Interval", 
     ylab="Average number of steps", 
     main="Average number of steps for intervals")


#Which 5-minute interval, on average across all the 
#days in the dataset, contains the maximum number 
#of steps?

daily_activity[which.max(daily_activity$mean), ]$interval
#835
```

5 The 5-minute interval that, on average, contains the maximum number of steps
```{r echo = TRUE}
# Imputing missing values

#Calculate and report the total number of missing 
#values in the dataset (i.e. the total number of 
#rows with \color{red}{\verb|NA|}NAs)
sum(is.na(activity$steps))
```
6 Code to describe and show a strategy for imputing missing data

```{r echo = TRUE}

#Devise a strategy for filling in all of the missing 
#values in the dataset. The strategy does not need to 
#be sophisticated. For example, you could use the
#mean/median for that day, or the mean 
#for that 5-minute interval, etc.
newdevise_steps <- daily_activity$mean[match(activity$interval, daily_activity$interval)]

#Create a new dataset that is equal to the original 
#dataset but with the missing data filled in.
 
new_dataset <- transform(activity, steps = ifelse(is.na(activity$steps), yes = newdevise_steps, no = activity$steps))
total_steps_new <- aggregate(steps ~ date, new_dataset, sum)
names(total_steps_new) <- c("date", "daily_steps")


```



7 Histogram of the total number of steps taken each day after missing values are imputed

```{r echo = TRUE}

#Make a histogram of the total number of steps taken 
# day and Calculate and report the mean and median
#total number of steps taken per day. Do these values 
#differ from the estimates from the first part of the 
#assignment? What is the impact of imputing missing 
#data on the estimates of the total daily number of steps?
hist(total_steps_new$daily_steps, col = "darkred", 
     xlab = "Total of steps for day", ylim = c(0,30), 
     main = "Total number of steps taken each day",
     breaks = seq(0,25000,by=2500))
#mean of the total number of steps taken per day
mean(total_steps_new$daily_steps)
#10766.19

#median of the total number of steps taken per day:
median(total_steps_new$daily_steps)
#10766.19

#Do these values differ from the estimates from the 
#first part of the assignment? 
#RES: YES ARE DIFFERENT

#What is the impact of imputing missing data on the 
#estimates of the total daily number of steps?
#RES: The mean and median values have increased and
#also have the same value.

```
8 Panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends
```{r echo = TRUE}
#Create a new factor variable in the dataset with two
#levels – “weekday” and “weekend” indicating whether
#a given date is a weekday or weekend day.

activity$date <- as.Date(strptime(activity$date, format="%Y-%m-%d"))
activity$datetype <- sapply(activity$date, function(x) {
  if (weekdays(x) == "sábado" | weekdays(x) =="domingo") 
  {y <- "Weekend"} else 
  {y <- "Weekday"}
  y
})

 
```
9 All of the R code needed to reproduce the results (numbers, plots, etc.) in the report

```{r echo = TRUE}

#Make a panel plot containing a time series plot 
#(i.e. \color{red}{\verb|type = "l"|}type = "l") 
#of the 5-minute interval (x-axis) and the average
#number of steps taken, averaged across all weekday 
#days or weekend days (y-axis). See the README file
#in the GitHub repository to see an example of what
#this plot should look like using simulated data.

activity_date <- aggregate(steps~interval + datetype, activity, mean, na.rm = TRUE)
plot<- ggplot(activity_date, aes(x = interval , y = steps, color = datetype)) +
    geom_line() +
    labs(title = "Average daily steps for date", x = "Interval", y = "Average for number of steps") +
    facet_wrap(~datetype, ncol = 1, nrow=2)
  print(plot)

```









